% !TeX root = main.tex
\section{THEORETICAL FRAMEWORK} \label{sec:frameworks}
This work adopts theoretical framings from personality psychology, criminology, and education. This section briefly explains the concepts from these theoretical frameworks for the readers' benefit (In the next section, we explain how they relate to computing education and this work in particular). 

\subsection{Personality Psychology}
The Big Five personality traits, the Five Factor Model (FFM), represent a widely accepted framework for understanding and categorizing human personality. This model emerged due to extensive psychological research to identify the fundamental dimensions underlying individual personality differences. The development of the Big Five model began in the late 20th century, with early research conducted by psychologists such as Tupes and Christal in the 1960s \cite{tupesRECURRENTPERSONALITYFACTORS}, followed by the work of Goldberg in the 1980s and 1990s \cite{Goldberg1990, Goldberg}, which laid the groundwork for the modern understanding of these traits.

The Big Five traits encompass five broad dimensions, each representing a distinct aspect of personality:

\begin{enumerate}
    \item \textbf{Openness to Experience (O):} This trait reflects one's inclination towards intellectual curiosity, creativity, and openness to new ideas, experiences, and perspectives. Individuals high in openness tend to be imaginative, curious, and open-minded, while those low in openness may be more conventional, cautious, and resistant to change.
    \item \textbf{Conscientiousness (C):} Conscientiousness pertains to the degree of organization, responsibility, diligence, and self-discipline an individual exhibits. High conscientiousness is associated with reliability, thoroughness, and goal-directed behavior, whereas low conscientiousness may manifest as impulsivity, disorganization, and a lack of follow-through.
    \item \textbf{Extraversion (E):} Extraversion refers to the extent to which an individual is outgoing, sociable, assertive, and energetic in social interactions. High extraversion is characterized by traits such as sociability, talkativeness, and enthusiasm, while introversion is marked by a preference for solitude, introspection, and quieter activities.
    \item \textbf{Agreeableness (A):} Agreeableness encompasses kindness, empathy, cooperativeness, and compassion towards others. Individuals high in agreeableness tend to be altruistic, trusting, and accommodating, whereas those low in agreeableness may exhibit more competitive, skeptical, or antagonistic tendencies.
    \item \textbf{Neuroticism (N):} Neuroticism reflects the tendency to experience negative emotions such as anxiety, depression, irritability, and vulnerability to stress. High neuroticism is associated with traits such as emotional volatility, insecurity, and sensitivity to perceived threats, while low neuroticism is characterized by emotional resilience, calmness, and emotional stability.
\end{enumerate}

While certain researchers have raised concerns about the Big Five model, suggesting that its emphasis on just five overarching dimensions offers an incomplete depiction of personality traits \cite{Block1995ACV}, it remains extensively embraced and applied in psychological research. Moreover, modifications have been made to the number of questions utilized to derive the scores for these five dimensions  \cite{JOHNSON201478} making it more trustworthy and accurate. 

\subsection{Criminology}
The fraud triangle, a seminal concept in criminology, was introduced by sociologist and criminologist Donald R. Cressey in 1953 \cite{Cressey}. This framework suggests that three key factors - \textit{opportunity, motivation (or pressure), and rationalization} are typically present in occupational fraud or embezzlement. According to Cressey, individuals are more likely to engage in fraudulent behavior when they perceive an opportunity to commit the act, experience pressure or motivation to do so (often due to financial difficulties or other personal stressors), and can justify or rationalize their actions. 

The comparison between fraud and plagiarism, as well as the potential use of the Fraud Triangle concept in educational settings, has been already explored in existing literature \cite{fraudTriangleToAcadmics, FraudTriEffect}. It's important to note that such discussions don't equate students with criminals but rather examine parallels in behavior and circumstances.

%\subsection{Academic Integrity}
%Among these, plagiarism, defined as using others' work without proper acknowledgment, has received extensive scholarly attention.  Additionally, collaboration on assignments violating individual work requirements, termed \textit{collusion} \cite{JonesJuliet}, has emerged as a source of ambiguity concerning the distinction between acceptable conduct and academic impropriety.

\section{Related Work} \label{sec:relatedwork}
In this section, we explore the existing literature and the application of the frameworks introduced in the previous section.

\subsection{The Big Five Personality Traits in Academic Plagiarism}

Interest in the influence of the big five personality traits on academic plagiarism has grown in recent years. However, research in this area remains limited within computing education. Contradictory findings exist in the literature; for example, Bhutto et al. \cite{Bhutto2019ACS} explored the correlation between personality traits and plagiarism among 231 undergraduate-level students from social science departments, revealing \textit{positive} associations with agreeableness, conscientiousness, extraversion, and openness to experience, while no significant relationship was found with neuroticism. Conversely, Correa \cite{ChileanUniver} investigated cyber plagiarism among 106 Chilean undergraduate students across departments, finding a significant \textit{negative} correlation with conscientiousness, extraversion, and openness to experience and a positive correlation with neuroticism. Both of their works use a questionnaire measuring personality traits and self-reported plagiarism behavior.

Wilks et al. \cite{Wilks2016-WILPTA-3} conducted the same study in the Portuguese context with undergraduate students from Law and Criminology departments and found that Conscientiousness and Agreeableness traits are negatively correlated with the inclination to plagiarize, while no significant association was found with Neuroticism. 

Giluk et al.\cite{Giluk2015BigFP} performed a meta-analysis to estimate the relationship between each of the Big Five personality factors and academic dishonesty. Their findings indicate that conscientiousness and agreeableness are the strongest Big Five predictors, with both factors negatively related to academic dishonesty. Table \ref{tab:litSUmmary} summarizes the literature regarding the impact of each of the big five traits on academic plagiarism. A positive (+ve) or negative (-ve) sign denotes the direction of influence, while "-" indicates no significant effect. "NA" indicates not applicable (as it was a meta-analysis). 

\vspace{-8pt}
\begin{table}[H]
  \centering
  \caption{Summary of Big Five traits and Plagiarism\\\label{tab:litSUmmary}}
    \vspace{-12pt}
  \begin{tabular}{p{2cm}rrrrrr}
    \toprule
    Study & Participants & O & C & E & A & N \\\midrule
    Bhutto et al.\cite{Bhutto2019ACS} & 231 & +ve & +ve & +ve & +ve & - \\
    Correa\cite{ChileanUniver} & 106 & -ve & -ve & -ve & - & +ve \\
    Wilks et al.\cite{Wilks2016-WILPTA-3} & 373 & - & -ve & - & -ve & - \\
    Giluk et al.\cite{Giluk2015BigFP} & NA & - & -ve & - & -ve & - \\ \bottomrule
  \end{tabular}
  \vspace{-8pt}
\end{table}

The findings suggest inconclusiveness, necessitating further replication and substantiation through additional evidence. Moreover, it's crucial to recognize that these studies depend on self-reported occurrences of plagiarism or student collusion, which may not offer a precise representation due to the potential influence of fear regarding disciplinary consequences or social desirability bias. Notably, there exists a significant dearth of such research within computing education. Our study aims to fill this gap by addressing the inquiry outlined in our research question RQ1. Additionally, rather than relying on self-reported plagiarism behavior by students, we assess actual assignments to determine collusion and plagiarism scores.

\subsection{Plagiarism in Programming Assignments}
Plagiarism in computing education has been researched heavily in the past two decades. Naaj et al. \cite{8937362} explored the factors impacting plagiarism and collusion in programming assignments, revealing that female students and those with a higher CGPA tended to exhibit stronger ethical behavior compared to their peers. Additionally, no significant differences were observed based on student level or major. 

Karnalim et al \cite{9678917} conducted a study in Indonesia aimed at understanding student perceptions of programming plagiarism and collusion. Their survey of 345 students from 16 universities revealed that students generally recognize instances of programming plagiarism and collusion, except for behaviors like self-plagiarism, copying from early drafts of their own work, and seeking assistance to correct code. 

Karnalim et al \cite{10.1145/3506717} create and evaluate an assessment submission system with automated, personalized feedback to deter collusion and plagiarism. The system generates personalized similarity reports for submissions showing similarities, prompting students to clarify and resubmit their work. Otherwise, similarity reports are sent only to the submitting student to improve their understanding. Their research suggests a decrease in programming plagiarism and collusion following the system's implementation, although student engagement was lower. To address this, gamification elements were introduced, offering points and badges for reviewing generated reports. Those with the highest points were offered rewards. As a result, students demonstrated increased awareness of programming plagiarism and collusion, completed assessments sooner, and engaged more frequently with the reports \cite{10042043}.

Honor pledges have demonstrated effectiveness in educating students about plagiarism and collusion. Gibbs et al. \cite{10.1145/2724660.2728663} found that while honor codes had a non-significant effect, pre-task warnings significantly reduced cheating rates. Bates et al.\cite{bates2016academic} conducted experiments using two approaches to educate students about plagiarism: providing a student code of conduct and presenting explicit definitions of plagiarism from their university's senate rules. Both interventions resulted in shifts in students' perceptions of acceptable academic behavior, underscoring the importance of clearly defining boundaries for unacceptable conduct.

Albluwi \cite{10.1145/3371156} conducted a comprehensive systematic review of plagiarism in programming assignments in 2019 with 87 published papers from the lens of the fraud triangle. The review revealed that a majority (68\%) of the examined papers focused on methods to reduce \textit{opportunities} for plagiarism and tools for detecting it. However, there is a notable absence of empirical research assessing the effectiveness of these strategies and tools as deterrents. Additionally, the papers (33\%) discussed various \textit{rationalizations} employed by computing students to justify plagiarism, with genuine confusion about plagiarism definitions being a prominent factor. Additionally, research on the correlation between academic \textit{pressure} in computing courses and plagiarism was found to be limited.

Research studies have consistently highlighted significant confusion among students regarding the distinction between acceptable collaboration and unacceptable collusion \cite{10.1145/2591708.2591755, 10.1145/2632320.2632342, Joy2011SourceCP}. While researchers have underscored the importance of clearly communicating what constitutes plagiarism to students \cite{10.1145/1595496.1562900, 10.1145/3160489.3160502, 10.1145/1734263.1734365, 10.1145/3024906.3024910}, examining the effectiveness of such efforts is limited. This study addresses this gap through research question RQ2.


